Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-10-15 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:10:15	16:37:19	(Thu15.Oct)

In many cases MVDR can improve the contrast and resolution of active sonar images compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, while the computational complexity of conventional beamformers are linear with the number of channels, O($M$), MVDR is at O($M^3$). This is because MVDR relies on estimating and inverting a covariance matrix. We have shown that in systems with less than $\approx$32 channels the complexity can be reduced and the implementation accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014}. For larger systems the inversion step is the bottleneck and space reduction techniques are desirable. In clinical ultrasound, where the transmit beams are focused, we adapted our GPU implementation to operate in beamspace where the spatial sparsity allows much smaller matrices to be inverted~\cite{Asen2013}. Two closely related dimensionality reducing techniques used with MVDR includes using principal component analysis (PCA)~\cite{Kim2014}, and to assume spatial stationarity to form more easily invertible Toepliz matrices~\cite{Asl2012}. At best the mentioned methods are able to reduce the complexity to O($M^2$).

A much faster and simpler alternative to MVDR is the Low Complexity Adaptive (LCA) beamformer. Based on an idea by Vignon~\cite{Vignon2008}, it was first introduced by Synnev\aa{}g in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar image quality to MVDR in systems with focused transmit beams\todo{ha med at i US \o{}nsker man ulike vindus-typer p\aa{} forskjellig range?}. LCA applies a set of predefined windows, or tapers, and selects the one offering the best sidelobe suppression. It relies on the same optimization criterion as MVDR, thus may be considered a version of MVDR with a reduced window solution space. Alternatively, it may be viewed as a an multi-taper technique. One such method is described by Stankwitz in radar imaging~\cite{Stankwitz1995}. Here the best of 3 windows are selected, but no micro-steered variations are used like LCA does.

Synnev\aa{}g used a window set comprised of rectangular, Kaiser (or Kaiser-Harris) and inverted Kaiser functions. In total he made LCA choose from 12 unique windows, 6 of which were micro-steered variants of a fairly wide Kaiser window. While achieving good results, it was not made clear how LCA performs with more or fewer windows, and how they should be steered. Also, it is not clear how LCA performs in systems with unfocused transmit beams.

@@

In the image reconstruction we have run MVDR with subarray length $L=16$ and $\text{\large$\epsilon$}=1\%$ diagonal loading. Both MVDR and LCA was run with $K=1$ temporal averaging. This is a fairly aggressive yet stable set of parameters.\todo{suitable ref?}

@@

When using LCA with Kaiser windows we made their responses similar to those typically chosen by a robust MVDR in key areas in the image. We suggest an upper bound for $\beta$ between 2 and 5, where 5 is slightly better at suppressing sidelobes. For $\phi$ we suggest to steer either to 50\% or 100\% of the \minus{}3\,dB width, which leads to a resolution of 73\% or 54\% of the rectangular window, respectively. As lower boundaries we set $\beta=0$ to always include a rectangular window, and $\phi=0$ to always include the centered window.

>>>>

In many cases MVDR can improve the contrast and resolution of active sonar images compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, while the computational complexity of conventional beamformers are linear with the number of channels, O($M$), MVDR is at O($M^3$). This is because MVDR relies on estimating and inverting a covariance matrix. In sonar systems with less than 32 channels the complexity can be reduced and the implementation accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014}. For larger systems the inversion step dominates the computation, the use of approximations through space reduction techniques become attractive. Some examples of such techniques include beamspace proessing on a GPU~\cite{Asen2013}, the closely related principal component analysis (PCA) method~\cite{Kim2014}, or to assume spatial stationarity to form more easily invertible Toepliz matrices~\cite{Asl2012}. At best the mentioned methods are able to reduce the complexity to O($M^2$).

A much faster and simpler alternative to MVDR is the Low Complexity Adaptive (LCA) beamformer. Based on an idea by Vignon~\cite{Vignon2008}, it was first introduced by Synnev\aa{}g \emph{et al}. in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar image quality to MVDR in systems with focused transmit beams\todo{ha med at i US \o{}nsker man ulike vindus-typer p\aa{} forskjellig range?}. LCA applies a set of predefined windows, or tapers, and selects the one offering the best sidelobe suppression. It relies on the same optimization criterion as MVDR, thus may be considered a version of MVDR with a reduced window solution space. Alternatively, it may be viewed as a an multi-taper technique. One such method is described by Stankwitz \emph{et al}. in radar imaging~\cite{Stankwitz1995}. Here the best of 3 windows are selected, but unlike with LCA no micro-steered variations are used.

Synnev\aa{}g \emph{et al}. suggested using a window set comprised of rectangular, Kaiser (or Kaiser-Harris) and inverted Kaiser functions. In total he made LCA choose from 12 unique windows, 6 of which were micro-steered variants of a fairly wide Kaiser window. While achieving good results, it was not made clear how LCA performs with more or fewer windows, and how they should be steered. Also, it is not clear how LCA performs in systems with unfocused transmit beams.

@@

In the image reconstruction we have run MVDR with subarray length $L=16$ and $\text{\large$\epsilon$}=1\%$ diagonal loading. Both MVDR and LCA was run with $K=1$ temporal averaging. This is a fairly aggressive yet stable set of parameters~\cite{Synnevag2009a}.

@@

When using LCA with Kaiser windows we made their responses similar to those typically chosen by a robust MVDR in key areas in the image. We suggest an upper bound for $\beta$ between 2 and 5, where 5 is slightly better at suppressing sidelobes. For $\phi$ we suggest to steer either to 50\% or 100\% of the \minus{}3\,dB width, which leads to a resolution of 73\% or 53\% of the rectangular window, respectively. As lower boundaries we set $\beta=0$ to always include a rectangular window, and $\phi=0$ to always include the centered window.
